WGS fastq2vcf walkthrough

### 2022-02-04 WGS-Full-Walkthrough.Rmd

#### Example Fastq File from Sequencing Facility

##### Read 1
```
[schaal.s@login-00 CodGenomes]$ zcat i5-2-i7-9_R1_001.fastq.gz | head -n 10

@GWNJ-1012:218:GW191226406th:1:1101:1090:1000 1:N:0:GATCAG+ATAGAGGC
ACTTGACTGTGCGTTGGCCTGCGGGCTGACTCGGTCCTGAGATGGACTGCTGTGTAGTTTGAACCATAGATTCATTATATAGAACACGGTCTCCTCTGCGCTGCTGGCCAATGGAGCCGAACGTCCGCACTGGCGGGCGGCCATCTTGCC
+
FF:,FF,FFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFFFFFFFF,:FFFFFFFFF:FFFFFFF:FFFFFFFFFFFFF:FFFFFFF
@GWNJ-1012:218:GW191226406th:1:1101:1687:1000 1:N:0:GATCAG+ATAGAGGC
ACTTGATCCCTCTCACTCTCCTCTCCGTCTCCTCTTTTGTCCTCGTCTCTCTCCTCTCTCCCTCTCTCCCATCTCCCTCTCTATCAAGTAGATCGGAAGAGCACACGTCTGAACTCCAGTCACGATCAGATCTCGTATGCCGTCTTCTGC
+
FFFFFF,FFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFFFFFFFFF:FFFF:::F,F::FFFFF
@GWNJ-1012:218:GW191226406th:1:1101:6840:1000 1:N:0:GATCAG+ATAGAGGC
ACTTGAAAAAAATACATAGCGGCCATGGACAGGATGACCTCTATGACAATGATAGAAACAGAAAGGACGCGGAGACTCTTGAGTCATCAAGTAGATCGGAAGAGCACACGTCTGAACTCCAGTCACGATCAGATCTCGTATGCCGTCTTC
```	
##### Read 2
```
[schaal.s@login-00 CodGenomes]$ zcat i5-2-i7-9_R2_001.fastq.gz | head -n 10

@GWNJ-1012:218:GW191226406th:1:1101:1090:1000 2:N:0:GATCAG+ATAGAGGC
ACTTGAAGGCAAGATGGCCGCCCGCCAGTGCGGACGTTCGGCTCCATTGGCCAGCAGCGCAGAGGAGACCGTGTTCTATATAATGAATCTATGGTTCAAACTACACAGCAGTCCATCTCAGGACCGAGTCAGCCCGCAGGCCAACGCACA
+
FFFFFF,FFFF:FFFFFF:FFFFFFFFFFFFFFFFFF:FFFFFFF:,FFFFF:FF,FFFFFFFFFFFFFFF,F,:FFFF:F,FFFFF:FF,FFFF,FFFFFF,FFF,FF:FFFFF:F:FFFF,FFF:F:F:FFFFFFFFFFF,:FFFFFF
@GWNJ-1012:218:GW191226406th:1:1101:1687:1000 2:N:0:GATCAG+ATAGAGGC
ACTTGAAAGAGAGGGAGATGGGAGAGAGGGAGAGAGGAGAGAGACGAGGACAAAAGAGGAGACGGAGAGGAGAGTGAGAGGGATCAAGTAGATCGGAAGAGCGTCGTGTAGGGAAAGAGTGTGCCTCTATGTGTAGATCTCGGTGGTCGC
+
FFFFFF,FFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFF,FFFF:FFFFFFFFFFFFFFFFFFFFFFFFFFFF:FF,FF:
@GWNJ-1012:218:GW191226406th:1:1101:6840:1000 2:N:0:GATCAG+ATAGAGGC
ACTTGAAGACTCAAGAGTCTCCGCGTCCTTTCTGTTTCTATCATTGTCATAGAGGTCATCCTGTCCATGGCCGCTATGTATTTTTATCAAGTAGATCGGAAGAGCGTCGTGTAGGGAAAGAGTGTGCCTCTATGTGTAGATCTCGGTGGT
```

@@

Barcode File: For each set of paired end files that need to be demultiplexed, a barcode file needs to be created. For example, for files i5-4-i7-11_R1_001.fastq.gz and i5-4-i7-11_R2_001.fastq.gz its barcode file is as follows:

@@

This file contains the 12 barcodes that have the i5_4 and i7_11 index and which sample that combination belongs to. This file is tab delimited and because I had inline barcodes, meaning they were the same on both ends of the read, I have the same barcode sequence in the first and second column. Then the sample ID is on the right.

@@

#### Code for running Stacks
```
process_radtags -P -p {Path to genome files} -o .{Path for output} -b {barcode file} --inline_inline --disable_rad_check -r

@@

#### Example code for running on a Cluster 
```
#!/bin/bash
#SBATCH --job-name=stacks_i54_i711                # sets the job name
#SBATCH --mem=1Gb                                 # reserves 10 GB memory
#SBATCH --partition=lotterhos                     # requests that the job is executed in lotterhos partition 
#SBATCH --mail-user=schaal.s@northeastern.edu
#SBATCH --mail-type=FAIL
#SBATCH --time=24:00:00                           # reserves machines/cores for 24 hours.
#SBATCH --output=stacksi54_i711.%j.out                # sets the standard output to be stored in file, where %j is the job id
#SBATCH --error=stacksi54_i711.%j.err                 # sets the standard error to be stored in file

>>>>

#### Example Fastq File from Sequencing Facility 
To inspect an example data file, download files i5-4-i7-11_R1_exampleData.fastq.gz & i5-4-i7-11_R2_exampleData.fastq.gz which are truncated fastq files as examples for how data look from the sequencing facility.

##### Read 1
```
[schaal.s@login-00 workingGroupFiles]$ head i5-4-i7-11_R1_exampleData.fastq.gz 
@GWNJ-1012:218:GW191226406th:1:1101:1181:1000 1:N:0:GGCTAC+GGCTCTGA
CGATGTTTGGAAAGACAAAAAATACATTTAGATCCAAATGACTTCTATAATGGATTTTGAAATTATATATACGAGATGCTTTAGTTCAAAACATTTGACACAAATGAGGTTCAACATAAAGAACATCGAGATCGGAAGAGCACACGTCTG
+
FFF:FF,FFFFFFFFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFF:FFFFFFFFFF:F,FFF:FFFFFFFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFFFFFFFFFFFFFFFF,FFFFFFFFFFFFF:FFFF,F:FFFFFFF:FFF
@GWNJ-1012:218:GW191226406th:1:1101:1651:1000 1:N:0:GGCTAC+GGCTCTGA
CTTGCATTGAGATTTCTGAAAGAAGAGCTCCATCCAGTGGCAGAAAGGGGACGTTTCAAAAGCACGAATGCAAGCCTTTTAATTTTTCCAAAATAAAAATATGCAAGAGATCGGAAGAGCACACGTCTGAACTCCAGTCACGGCTACATC
+
FFFFFF,FFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFF
@GWNJ-1012:218:GW191226406th:1:1101:2193:1000 1:N:0:GGCTAC+GGCTCTGA
CGATGTTGCATGGAGTGGCTGGTGTCTGTGCCGACCACTCTGGTTCCAGAGGTAATGAAGACCAGCTTCCCCCCTGGCCACACAGTGGACTCAATCTTCCCTGAGATAAAGGTGAAGAGTTCCTCCTTGTTCTTAGGGTAACATCGAGAT
```	
##### Read 2
```
[schaal.s@login-00 workingGroupFiles]$ head i5-4-i7-11_R2_exampleData.fastq.gz 
@GWNJ-1012:218:GW191226406th:1:1101:1181:1000 2:N:0:GGCTAC+GGCTCTGA
CGATGTGCTTTATGTTGAACCTCATTTGTGTCAAATGTTTTGAACTAAAGCATCTCGTATATATAATTTCAAAATCCATTATAGAAGTCATTTGGATCTAAATGTATTTTTTGTCTTTCCAAACATCGAGATCGGAAGAGCGTCGGGTAG
+
FFFFFF,FF,FFFFFFFF:FFFFFFFFFFFFFFFFFFFFFFF:FFFFFFFFF:FFFF:F,F:FFF,:FFFFFFFFFF,,:,:FFFFFFFF,FFFFFFFFF:F:FFF:F,F,FFFFF:FFF:FFF,,F::FFFFF,,F:FFFFFFF,F,,F
@GWNJ-1012:218:GW191226406th:1:1101:1651:1000 2:N:0:GGCTAC+GGCTCTGA
CTTGCATATTTTTATTTTGGAAAAATTAAAAGGCTTGCATTCGTGCTTTTGAAACGTCCCCTTTCTGCCACTGGATGGAGCTCTTCTTTCAGAAATCTCAATGCAAGAGATCGGAAGAGCGTCGTGTAGGGAAAGAGTGTTCAGAGCCGT
+
FFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFFFF:FFFFFFFFFFFFFFF,
@GWNJ-1012:218:GW191226406th:1:1101:2193:1000 2:N:0:GGCTAC+GGCTCTGA
CGATGTTACCCTAAGAACAAGGAGGAACTCTTCACCTTTATCTCAGGGAAGATTGAGTCCACTGTGTGGCCAGGGGGGAAGCTGGTCTTCATTACCTCTGGAACCAGAGTGGTCGGCACAGACACCAGCCACTCCATGCAACATCGAGAT
```

What these files show are XXXX.

@@

Barcode File: For each set of paired end files that need to be demultiplexed, a barcode file needs to be created. For example, for files i5-4-i7-11_R1_exampleData.fastq.gz and i5-4-i7-11_R2_exampleData.fastq.gz its barcode file is as follows:

@@

This file contains the 12 barcodes that have the i5_4 and i7_11 index and which sample that combination belongs to. This file is tab delimited and because I had inline barcodes, meaning they were the same on both ends of the read, I have the same barcode sequence in the first and second column. Then the sample ID is in the far right column.

@@

#### Code for running Stacks
```
process_radtags -P -p {Path to genome files} -o {Path for output} -b {barcode file} --inline_inline --disable_rad_check -r

@@

#### Example code for running on a Cluster 
```
#!/bin/bash
#SBATCH --job-name=stacks_i54_i711                # sets the job name
#SBATCH --mem=1Gb                                 # reserves 1 GB memory
#SBATCH --mail-user=schaal.s@northeastern.edu     # add your email address to get notifications
#SBATCH --mail-type=FAIL                          # set what email notifications you want to receive here I only want emails when my script fails
#SBATCH --time=24:00:00                           # reserves machines/cores for 24 hours.
#SBATCH --output=stacksi54_i711.%j.out            # sets the standard output to be stored in file, where %j is the job id
#SBATCH --error=stacksi54_i711.%j.err             # sets the standard error to be stored in file
